---
title: "Study case"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message = FALSE, warning = FALSE, fig.height=8, fig.width=11)
library(knitr)
library(tidyverse)
library(BiocManager)
library(scuttle)
library(scater)
library(uwot)
library(edgeR)
library(ruv)
library(Seurat)
library(swapper)
library(cluster)
library(factoextra)
library(UpSetR)
library(foreach)
library(msigdbr)


source("/mnt/auxf.R")


#ds <- readRDS("/domino/datasets/local/RUV/sclupusds4.rds")
PBC <- readRDS("/domino/datasets/local/RUV/pblupusds4.rds")
path <- ('/domino/datasets/local/RUV')


```

```{r DGE Processing cohorts}

DiffGPCs <- readRDS("/domino/datasets/local/RUV/DiffGPCs_popE.rds")

topallDPC <- lapply(DiffGPCs, function (x) dplyr::arrange(x,P.Value))
#logFC_PC <- lapply(topallDPC, function(x) dplyr::select(x,logFC))

DiffPC.sig<- lapply(DiffGPCs, function (x) dplyr::filter(x,adj.P.Val<0.01))
topDPC <- lapply(DiffPC.sig, function(x) arrange(x,P.Value) )
#logFC_PC <- lapply(topDPC, function(x) x$logFC)


```

```{r DGE Processing cohorts alt code, include=FALSE}
# Alternatively

# ds2 <- readRDS("/domino/datasets/local/RUV/sclupusds2.rds")
# ds2 <- ds2[,ds2$Sex=='Female'] # remove male samples
# Idents(object = ds2) <- "cg_cov"
# celltypes <- c('B', 'NK', 'T4', 'T8', 'cDC', 'cM', 'ncM', 'pDC')
# ds2s <- subset(x = ds2, idents = celltypes)
# rm(list = 'ds2')
# 
# pdata <- droplevels(ds2s@meta.data)
# pdata$Age <- as.numeric(as.character(pdata$Age))
# 
# ds2s <- SetIdent(ds2s, value = "sample_cell") # change ident to group later
# ds2s <- FindVariableFeatures(ds2s) #### adds metafeatures on ds2s@assays[["originalexp"]]@meta.features
# ds2s <- ScaleData(ds2s) #### adds scaledata on ds2s@assays[["originalexp"]]@data
# 
# PBC <- AggregateExpression(ds2s, group_by = "sample_cell", assays = 'originalexp',slot='counts', return.seurat = F)
# PBC <- PBC$originalexp # same result than manual procedure
# colnames(PBC) <- unique(pdata$sample_cell)
# metacovs <- colnames(pdata)[c(4:6,8,10:16, 18:19)] # variables at type cell-sample level
# pbc.metaD <- unique(pdata[,metacovs])
# rownames(pbc.metaD) <- pbc.metaD$sample_cell
# pbc.metaD$Age <- as.numeric(as.character(pbc.metaD$Age))
# 
# PBC <- PBC[rowSums(PBC >= 5) >= 5,]
# gene.D <- data.frame(gene=rownames(PBC))
# rownames(gene.D) <- gene.D$gene
# 
# 
#  DEGPC24 <- foreach(ct = celltypes) %do% {
# 
#    pbc.ct <- PBC[,pbc.metaD$cg_cov==ct]
#    pbc.ct <-  pbc.ct[rowSums(pbc.ct >= 5) >= 5,]
#    gene.D.ct <- gene.D[ rownames(pbc.ct),]
# 
#    y.ct <- DGEList(counts=pbc.ct )
#    pmin.ct <- find_p(pbc.ct )
#    y.ct <- calcNormFactors(y.ct, method="upperquartile", p=pmin.ct)
#    nf.ct <- y.ct$samples$norm.factors
#    logy.ct <- edgeR::cpm(y.ct,log = T)
# 
#    design <- model.matrix(~ 0 + pbc.metaD$Processing_Cohort[pbc.metaD$cg_cov==ct] )
#    colnames(design) <- c('PC1','PC2', 'PC3', 'PC4')
#    v <- voom(y.ct, design, plot=F)
#    vfit <- lmFit(v, design)
# 
#    contrast.matrix <- makeContrasts(
#      PC2 - PC1, PC3 - PC1, PC4 - PC1, PC3 - PC2, PC4 - PC2, PC4 - PC3,
#      levels = colnames(design)
#    )
# 
#    # Fit contrasts
# 
#    vfit <- contrasts.fit(vfit, contrast.matrix)
#    efit <- eBayes(vfit)
#    topTable(efit, number=dim(y.ct)[1], coef= "PC4 - PC2", adjust="BH") # previously I was taking just the F statistic, but here I focused on the PC 2 and PC 4
# 
#  }
# 
#  names(DEGPC24)<- celltypes
# 
#  saveRDS(DEGPC24,paste0(path,'/DiffGPC24.rds'))



```

```{r Negative control genes}

hkGagnon <- read.csv(paste0(path,"/Genes_Gagnon.txt"), sep="")
hkGagnon <- unlist(hkGagnon)
filter.vst<-FindVariableFeatures(PBC,selection.method = 'vst')
high_varg <- which(filter.vst[hkGagnon,]$vst.variance.standardized>1.8)
hk.ind <- rownames(PBC) %in% hkGagnon[-high_varg]

hkScmerge <- read.csv(paste0(path,"/ncgSCmerge.txt"), sep="",header = F)
hkScmerge <- unlist(hkScmerge)
high_varg.sc <- which(filter.vst[hkScmerge,]$vst.variance.standardized>1.8)
hk.ind2 <- rownames(PBC) %in% hkGagnon[-high_varg.sc]

gene.D <- data.frame(gene=rownames(PBC))
rownames(gene.D) <- gene.D$gene
celltypes <- levels(PBC$samples$cg_cov)


```

```{r pbps read}

PBPSC <- readRDS( "/domino/datasets/local/RUV/pbps10rep_studycasemix.rds")


# Alternatively: 

# PBPSC <- gen_PBPS(ds=ds, PBC=PBC, ctype='cg_cov', BioVar=c('Status','cg_cov','pop_cov'), NVar='Processing_Cohort', id_pb='sample_cell', id_sub='ind_cov', cell_id='cell_id', n=10, Seed=2 )
# 
# write_rds(PBPSC, "/domino/datasets/local/RUV/pbps10rep_studycasemix.rds")

```

# Dataset

We have N = `r length(unique(PBC$samples$ind_cov_batch_cov))` single-cell samples from J = `r length(unique(PBC$samples$ind_cov))` subjects selected from the lupus dataset, distributed as follows: 

```{r subset}

group_by(PBC$samples,Processing_Cohort,ind_cov,pop_cov,Age,Status) %>%
  summarise(count=n()) %>% ungroup()%>% 
  mutate(ind_cov=str_extract(as.character(ind_cov),"[^_]+$")) %>%
  ggplot(aes(y=ind_cov,x=Processing_Cohort, group=ind_cov, color=Status, shape = pop_cov)) +
  geom_point(size=2) + geom_line() + scale_colour_brewer(palette='Set1') +
  labs(y = "Subject", x= "Processing cohort", color = "Status", shape = "Ethnicity")+ theme_minimal()+
  ggtitle('Samples')


```


The dataset has no technical replicates between the healthy controls of processing cohorts 2 and 4. There is also a confounding effect between processing cohorts and the disease status, observed for instance in the PCA of the CD4 T-cells


<!-- ## Metrics functions -->

```{r metrics}

split.ct <- function(PBC,ct){
  
  y.ct <- PBC[rowSums(PBC$counts >= 10) >= 5,PBC$samples$cg_cov==ct]
  pmin <- find_p(y.ct)
  y.ct <- calcNormFactors(y.ct, method="upperquartile", p=pmin)
  
return(y.ct)
}

PCAplot <- function (pca,y,avg_sw2=F,avg_swp=F){
  
  if(isTRUE(avg_sw2|avg_swp)){
   labelgg <- paste0(c('2 PCs: ','10 PCs: '), c(avg_sw2,avg_swp)) 
  }else{
   labelgg <- ""
  }
  
  if(is.null(rownames(pca))){
    rownames(pca)<- y$samples$sample_cell
  }
  
  pos_x <- min(pca[,1])
  pos_y <- min(pca[,2]) 
  ppca <- data.frame(pca) %>%
  tibble::rownames_to_column("sample_cell") %>% # Add the metadata into this data frame; match by sample IDs
  dplyr::inner_join(y$samples, by = "sample_cell") %>%
  ggplot(aes(x = PC1,y = PC2,color= Processing_Cohort)) + geom_point(aes(shape=Status),size=3) + 
    #ggtitle("Logcpm plot") +
    labs(color = "Processing cohort", shape = "Status") + 
    theme_minimal()+ theme(legend.position = "bottom") + scale_color_brewer(palette='Set1')+
    annotate("text", x = c(pos_x+3,pos_x+3), y = c(pos_y+3, pos_y), label = labelgg, size=3, fontface="bold")
    
  return(ppca)
  
}

Silplot <- function (sil){
  
  fviz_silhouette(sil)+ scale_fill_brewer(palette = 'Set1') + 
    scale_color_brewer(palette = 'Set1') + theme_minimal() +
    theme(axis.text.x=element_blank(),legend.position = "bottom") +
    labs(fill="Processing Cohort",color="Processing Cohort")
}

RLEplot <-function(names, logy, samples){
   
  median <- apply(logy, 1, median)
  samples <- as.data.frame(samples)
  
  rle <- apply(logy, 2, function(x) x - median)
  dataplot <- as.data.frame(rle) %>% 
    pivot_longer(everything(),names_to = 'sample_cell') %>% 
    left_join(samples, by='sample_cell')
  
  dataplot$sample_cell <- fct_relevel(dataplot$sample_cell,names)
  
  ggplot(dataplot,aes(x=sample_cell, y=value,colour = Processing_Cohort, linetype=pop_cov )) +
    geom_boxplot(outlier.shape = NA) + ylim(-2,2) + scale_color_brewer(palette='Set1') + ggtitle("RLE plot") +
     geom_hline(yintercept=0, linetype='dashed') + labs(y="Relative log expression", color='Processing cohort')+
    theme_minimal()+ theme( legend.position = 'bottom',strip.text =element_blank(),
                            axis.title.x = element_blank(),
                            axis.text.x = element_blank(),
                            axis.ticks.x = element_blank())

}

histpvals <- function(y,topDPC){
  
  design <- model.matrix(~ y$samples$Status ) 
  colnames(design) <- c('Intercept','Managed')
  v <- voom(y, design, plot=F) 
  vfit <- lmFit(v, design)
  efit <- eBayes(vfit)
  pvalsUQ <- topTable(efit, coef=2, number=dim(y)[1], adjust="BH")
  pvalsUQ <- arrange(pvalsUQ, P.Value)
  
  pos <- which(abs(pvalsUQ$adj.P.Val-0.01)==min(abs(pvalsUQ$adj.P.Val-0.01))) 
  q.val <- pvalsUQ$P.Value[pos[1]]
  vp <- rownames_to_column(pvalsUQ,'gene')%>% mutate(DEP=if_else(gene %in% rownames(topDPC),T,F)) %>% ggplot( aes(x=logFC,y=-log10(P.Value), color= DEP))+ geom_point(alpha=0.5)+ geom_hline(aes(yintercept=-log10(q.val)),colour='darkgrey' )+ scale_color_brewer(palette='Set1') + ggtitle("Volcano plot") + theme_minimal()
  p <- ggplot(pvalsUQ, aes(x=P.Value)) + geom_histogram(bins=15) + scale_color_brewer(palette='Set1') + ggtitle("P-values histogram") + theme_minimal()
  
  
  
  return(list(pvals=pvalsUQ,hist =p,volcano=vp))
  
}

logfcs <- function(DPC, pvals,names=c('disease status','UQ')){
  
  colnames(DPC) <- paste0(colnames(DPC),'_pc')
  
  datatoplot <- inner_join(rownames_to_column(DPC,'gene'),rownames_to_column(pvals,'gene'), by='gene')
  
  cormethods <- paste0('Cor: ',round(cor(datatoplot$logFC_pc,datatoplot$logFC),2))
  
  pos_x <- min(datatoplot$logFC_pc)
  pos_y <- min(datatoplot$logFC)
  
  ggplot(datatoplot, aes(x=logFC_pc, y=logFC))+geom_point()+ #scale_color_continuous(type = "viridis")+
    labs(x =paste0("LogFC ",names[1]), y = paste0("LogFC ",names[2])) + 
    theme_minimal()+ theme(legend.position = "bottom") + ggtitle('Log Fold Changes')+
    annotate("label", x = pos_x+1, y = pos_y, label = cormethods, size=3, fontface="bold")

  
  
}

stvar <- function (logy){
  
 top_high_varg.ct <- FindVariableFeatures(logy)
top_high_varg.ct <- rownames(arrange(top_high_varg.ct,-vst.variance.standardized))[1:500]
return(top_high_varg.ct)
}




```



## Unwanted variation in the data {.tabset}


```{r raw data, results='hide'}

ys.ct <- lapply(celltypes,split.ct,PBC=PBC)

logys.ct <- lapply(ys.ct,function (x) voom(x,plot=F)$E)
  
hvg <- lapply(logys.ct,stvar)
pcas <- mapply(calculatePCA, x= logys.ct, ncomponents=rep(10,length(logys.ct)), subset_row=hvg, SIMPLIFY = F)

pcos <- lapply(ys.ct, function (x) as.integer(x$samples$Processing_Cohort))
# pcos <- lapply(pcos, function (x) {x[x==3]<-4; x})
sil2s <- mapply(silhouette,x=pcos,dist=lapply(pcas,function(x) dist(x[,1:2],"euclidean") ), SIMPLIFY = F)
silps <- mapply(silhouette,x=pcos,dist=lapply(pcas,dist, "manhattan") ,SIMPLIFY = F)

avgs_sw2 <- lapply(sil2s, function (x) round(mean(x[,3]),2))
avgs_swp <- lapply(silps, function (x) round(mean(x[,3]),2))
  
ordnames <- lapply(ys.ct, function (x) arrange(x$samples,Processing_Cohort)$'sample_cell') 

Silplots <- lapply(silps,Silplot)

PCAplots <- mapply(PCAplot, pca= pcas, y=ys.ct, SIMPLIFY = F)

RLEplots <- mapply(RLEplot, names= ordnames, logy=logys.ct,samples=map(ys.ct, "samples"), SIMPLIFY = F)


```


```{r rawpcaplots , results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n### ',celltypes[i], ' {-}\n')))
  
  plots <- list(PCAplots[[i]],Silplots[[i]], RLEplots[[i]])
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=1,ncol=3, common.legend = T, legend="bottom"))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```


## Naive DEA {.tabset}

Volcano Plot with differentially expressed genes across processing cohorts in blue

<!-- ### Code -->

```{r DEA Naive}

Hists <- mapply(histpvals,y=ys.ct,topDPC=topDPC,SIMPLIFY = F)


```


```{r DErawplots , results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n### ',celltypes[i], ' {-}\n')))
  
  plots <- list( Hists[[i]]$hist, Hists[[i]]$volcano 
                 )
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=1,ncol=2))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```


```{r export raw T4, include=FALSE}

PCAplot(pca= pcas[[3]], y=ys.ct[[3]], avg_sw2 = avgs_sw2[[3]] ,avg_swp = avgs_swp[[3]] ) 
ggsave(paste0(path,'/paper/graphs/studycase/T4PCArawSCm.png'),height = 5, width = 7)

RLEplots[[3]]
ggsave(paste0(path,'/paper/graphs/studycase/T4RLErawSCm.png'),height = 5, width = 7)

Hists[[3]]$hist
ggsave(paste0(path,'/paper/graphs/studycase/T4histrawSCm.png'),height = 5, width = 7)

```


## {-}

# Analysis


We will evaluate the PCA plots and average silhouette width (ASW) using the processing cohort as label, from the normalised matrices returned from the RUVIII and RUVIII PBPS methods, and the normalised matrix obtained by subtracting the processing cohort effect. We also include the p-value histogram and the volcano plot from the differential expression analysis between patients and controls.


## Including the processing cohort effect {.tabset}

<!-- ### Code -->

```{r pc, results = 'hide'}

normpc <- function(y){
  logy <- voom(y,plot=F)$E
  design <- model.matrix(~ y$samples$Status + y$samples$Processing_Cohort + y$samples$pop_cov + y$samples$Status*y$samples$pop_cov ) 
  #colnames(design) <- c('Intercept','tr',paste0('Proc',  c(2,4)))
  v <- voom(y, design, plot=F) 
  vfit <- lmFit(v, design)
  alpha <- vfit$coefficients[,3:5] # same coefficients as efit and returned by toptable
  newY <- logy - t(design[,c(3:5)]%*%t(alpha))
  return(newY)
}

histpc <- function(y,topDPC){
  logy <- voom(y,plot=F)$E
  design <- model.matrix(~ y$samples$Status + y$samples$Processing_Cohort + y$samples$pop_cov + y$samples$Status*y$samples$pop_cov ) 
  colnames(design) <- c('Intercept','tr',paste0('Proc',  c(2,3,4)),'pop','interaction')
  v <- voom(y, design, plot=F) 
  vfit <- lmFit(v, design)
  efit <- eBayes(vfit)
  pvalstr <- topTable(efit,coef=2, number=dim(y)[1], adjust="BH")
  pvalsint <- topTable(efit,coef=6, number=dim(y)[1], adjust="BH")
  contrast.matrix <- makeContrasts(
      SLEeuro =tr,
      SLEasian = interaction,
      levels = colnames(design))
    
    vfit <- contrasts.fit(vfit, contrast.matrix)
    efit <- eBayes(vfit)
  pvalspc <- topTable(efit, number=dim(y)[1], adjust="BH")
  
  pos <- which(abs(pvalstr$adj.P.Val-0.01)==min(abs(pvalstr$adj.P.Val-0.01))) 
  q.val <- pvalstr$P.Value[pos[1]]
  vptr <- rownames_to_column(pvalstr,'gene')%>% mutate(DEP=if_else(gene %in% rownames(topDPC),T,F)) %>% ggplot( aes(x=logFC,y=-log10(P.Value), color= DEP))+ geom_point(alpha=0.5)+ geom_hline(aes(yintercept=-log10(q.val)),colour='darkgrey' )+ scale_color_brewer(palette='Set1') + ggtitle("Volcano plot SLE") + theme_minimal()
  
  pos <- which(abs(pvalsint$adj.P.Val-0.01)==min(abs(pvalsint$adj.P.Val-0.01))) 
  q.val <- pvalsint$P.Value[pos[1]]
  vpint <- rownames_to_column(pvalsint,'gene')%>% mutate(DEP=if_else(gene %in% rownames(topDPC),T,F)) %>% ggplot( aes(x=logFC,y=-log10(P.Value), color= DEP))+ geom_point(alpha=0.5)+ geom_hline(aes(yintercept=-log10(q.val)),colour='darkgrey' )+ scale_color_brewer(palette='Set1') + ggtitle("Volcano plot SLE*Ethnicity") + theme_minimal()

  p <- ggplot(pvalspc, aes(x=P.Value)) + geom_histogram(bins=15) + scale_color_brewer(palette='Set1') + ggtitle("P-values histogram") + theme_minimal()
 
  return(list(pvals=pvalspc, pvalstr=pvalstr, pvalsint=pvalsint, hist =p, volcano=vptr, volcano_int=vpint))
  
  
  
}


normpcs.ct <- lapply(ys.ct,normpc)
hvg <- lapply(normpcs.ct,stvar)
pcas <- mapply(calculatePCA, x= normpcs.ct, ncomponents=rep(10,length(logys.ct)), subset_row=hvg, SIMPLIFY = F)
pcos <- lapply(ys.ct, function (x) as.integer(x$samples$Processing_Cohort))
# pcos <- lapply(pcos, function (x) {x[x==3]<-4; x})


sil2s <- mapply(silhouette,x=pcos,dist=lapply(pcas,function(x) dist(x[,1:2],"euclidean") ), SIMPLIFY = F)
silps <- mapply(silhouette,x=pcos,dist=lapply(pcas,dist, "manhattan") ,SIMPLIFY = F)

avgs_sw2 <- lapply(sil2s, function (x) round(mean(x[,3]),2))
avgs_swp <- lapply(silps, function (x) round(mean(x[,3]),2))
  
ordnames <- lapply(ys.ct, function (x) arrange(x$samples,Processing_Cohort)$'sample_cell') 

Silplots <- lapply(silps,Silplot)

PCAplots <- mapply(PCAplot, pca= pcas, y=ys.ct, SIMPLIFY = F)

RLEplots <- mapply(RLEplot, names= ordnames, logy=normpcs.ct, samples=map(ys.ct, "samples"), SIMPLIFY = F)

HistsB <- mapply(histpc, y=ys.ct,topDPC=topDPC,SIMPLIFY = F)
  

  
```

### Normalisation {.tabset}

```{r PCplots , results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n#### ',celltypes[i], ' {-}\n')))
  
  plots <- list( PCAplots[[i]],Silplots[[i]], RLEplots[[i]])
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=1,ncol=3, common.legend = T, legend="bottom"))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```

### DEA {.tabset}

```{r DEbatchplots , results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n#### ',celltypes[i], ' {-}\n')))
  
  plots <- list( HistsB[[i]]$hist,HistsB[[i]]$volcano,HistsB[[i]]$volcano_int)
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=2,ncol=2))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```


```{r export Batch T4, include=FALSE}

PCAplot(pca= pcas[[3]], y=ys.ct[[3]], avg_sw2 = avgs_sw2[[3]] ,avg_swp = avgs_swp[[3]] ) 
ggsave(paste0(path,'/paper/graphs/studycase/T4PCABatchSCm.png'),height = 5, width = 7)

RLEplots[[3]]
ggsave(paste0(path,'/paper/graphs/studycase/T4RLEBatchSCm.png'),height = 5, width = 7)

HistsB[[3]]$hist
ggsave(paste0(path,'/paper/graphs/studycase/T4histBatchSCm.png'),height = 5, width = 7)

```

## RUVIIIgb {.tabset}

<!-- ### Code -->
  
```{r ruv3, results='hide'}

histT <- function (y,w,topDPC,samples){
  k <- ncol(w)
  design <- model.matrix(~ samples$Status + w + samples$pop_cov + samples$Status*samples$pop_cov ) 
  colnames(design) <- c('Intercept','tr', paste0("W",1:k),'pop','interaction')
  v <- voom(y, design, plot=F) 
  vfit <- lmFit(v, design)
  efit <- eBayes(vfit)
  pvalstr <- topTable(efit,coef=2, number=dim(y)[1], adjust="BH")
  pvalsint <- topTable(efit,coef=6, number=dim(y)[1], adjust="BH")
  contrast.matrix <- makeContrasts(
      SLEeuro =tr,
      SLEasian = interaction,
      levels = colnames(design))
    
    vfit <- contrasts.fit(vfit, contrast.matrix)
    efit <- eBayes(vfit)
  pvalsT <- topTable(efit, number=dim(y)[1], adjust="BH")
  
  pos <- which(abs(pvalstr$adj.P.Val-0.01)==min(abs(pvalstr$adj.P.Val-0.01))) 
  q.val <- pvalstr$P.Value[pos[1]]
  vptr <- rownames_to_column(pvalstr,'gene')%>% mutate(DEP=if_else(gene %in% rownames(topDPC),T,F)) %>% ggplot( aes(x=logFC,y=-log10(P.Value), color= DEP))+ geom_point(alpha=0.5)+ geom_hline(aes(yintercept=-log10(q.val)),colour='darkgrey' )+ scale_color_brewer(palette='Set1') + ggtitle("Volcano plot SLE") + theme_minimal()
  
  pos <- which(abs(pvalsint$adj.P.Val-0.01)==min(abs(pvalsint$adj.P.Val-0.01))) 
  q.val <- pvalsint$P.Value[pos[1]]
  vpint <- rownames_to_column(pvalsint,'gene')%>% mutate(DEP=if_else(gene %in% rownames(topDPC),T,F)) %>% ggplot( aes(x=logFC,y=-log10(P.Value), color= DEP))+ geom_point(alpha=0.5)+ geom_hline(aes(yintercept=-log10(q.val)),colour='darkgrey' )+ scale_color_brewer(palette='Set1') + ggtitle("Volcano plot SLE*Ethnicity") + theme_minimal()

  p <- ggplot(pvalsT, aes(x=P.Value)) + geom_histogram(bins=15) + scale_color_brewer(palette='Set1') + ggtitle("P-values histogram") + theme_minimal()
  
  return(list(pvals=pvalsT, pvalstr=pvalstr, pvalsint=pvalsint, hist=p, volcano=vptr, volcano_int=vpint))
  
  
}

k=3
  
Mcts.ct <- lapply (ys.ct, function(x){
  Mct <- replicate.matrix(x$samples$ind_cov)
  rownames(Mct) <- x$samples$ind_cov
  Mct
})

ruv3T2 <- mapply(RUVIIIW, Y = lapply(logys.ct,t), M=Mcts.ct , ctl =  lapply(ys.ct, function (x) rownames(x) %in% hkGagnon[-high_varg]), k= as.list(rep(k,length(celltypes))), return.info=as.list(rep(T,length(celltypes))), SIMPLIFY=F)

normT2 <- lapply(ruv3T2,function (x) t(x$newY))

hvg <- lapply(normT2,stvar)
pcas <- mapply(calculatePCA, x= normT2, ncomponents=rep(10,length(logys.ct)), subset_row=hvg, SIMPLIFY = F)

sil2s <- mapply(silhouette,x=pcos,dist=lapply(pcas,function(x) dist(x[,1:2],"euclidean") ), SIMPLIFY = F)
silps <- mapply(silhouette,x=pcos,dist=lapply(pcas,dist, "manhattan") ,SIMPLIFY = F)

avgs_sw2 <- lapply(sil2s, function (x) round(mean(x[,3]),2))
avgs_swp <- lapply(silps, function (x) round(mean(x[,3]),2))

ordnames <- lapply(ys.ct, function (x) arrange(x$samples,Processing_Cohort)$'sample_cell') 

SilplotsT23 <- lapply(silps,Silplot)

PCAplotsT23 <- mapply(PCAplot, pca= pcas, y=ys.ct, SIMPLIFY = F)

RLEplotsT23 <- mapply(RLEplot, names= ordnames, logy=normT2,samples= map(ys.ct,'samples'), SIMPLIFY = F)

HistsT23 <- mapply(histT,y=ys.ct,w=map(ruv3T2,"W"), topDPC=topDPC, map(ys.ct,'samples'), SIMPLIFY = F)
  
#LFCplots <- mapply(function (x,y) logfcs(DPC=x, pvals=y$pvals,names=c('True','RUVIII')),x=trueLFC, y=HistsT23, SIMPLIFY=F)

 
```


### Normalisation {.tabset}


```{r ruv3plots gagnon , results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n#### ',celltypes[i], ' {-}\n')))
  
  plots <- list( PCAplotsT23[[i]],SilplotsT23[[i]], RLEplotsT23[[i]])
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=1,ncol=3, common.legend = T, legend="bottom"))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```

### DEA {.tabset}


```{r DE3plots gagnon, results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n#### ',celltypes[i], ' {-}\n')))
  
  plots <- list( HistsT23[[i]]$hist,HistsT23[[i]]$volcano,HistsT23[[i]]$volcano_int)
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=2,ncol=2))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```

```{r export ruv3 T4 gagnon, include=FALSE}

PCAplot(pca= pcas[[3]], y=ys.ct[[3]], avg_sw2 = avgs_sw2[[3]] ,avg_swp = avgs_swp[[3]] ) 
ggsave(paste0(path,'/paper/graphs/studycase/T4PCA3SCm.png'),height = 5, width = 7)

RLEplotsT23[[3]]
ggsave(paste0(path,'/paper/graphs/studycase/T4RLE3SCm.png'),height = 5, width = 7)

HistsT23[[3]]$hist
ggsave(paste0(path,'/paper/graphs/studycase/T4hist3SCm.png'),height = 5, width = 7)

```


## RUVIII scmerge {.tabset}

<!-- ### Code -->
  
```{r ruv3 scmerge, results='hide'}


ruv3T2scm <- mapply(RUVIIIW, Y = lapply(logys.ct,t), M=Mcts.ct , ctl =  lapply(ys.ct, function (x) rownames(x) %in% hkScmerge[-high_varg]), k= as.list(rep(k,length(celltypes))), return.info=as.list(rep(T,length(celltypes))), SIMPLIFY=F)

normT2scm <- lapply(ruv3T2scm,function (x) t(x$newY))

hvg <- lapply(normT2scm,stvar)
pcas <- mapply(calculatePCA, x= normT2scm, ncomponents=rep(10,length(logys.ct)), subset_row=hvg, SIMPLIFY = F)

sil2s <- mapply(silhouette,x=pcos,dist=lapply(pcas,function(x) dist(x[,1:2],"euclidean") ), SIMPLIFY = F)
silps <- mapply(silhouette,x=pcos,dist=lapply(pcas,dist, "manhattan") ,SIMPLIFY = F)

avgs_sw2 <- lapply(sil2s, function (x) round(mean(x[,3]),2))
avgs_swp <- lapply(silps, function (x) round(mean(x[,3]),2))

ordnames <- lapply(ys.ct, function (x) arrange(x$samples,Processing_Cohort)$'sample_cell') 

SilplotsT23scm <- lapply(silps,Silplot)

PCAplotsT23scm <- mapply(PCAplot, pca= pcas, y=ys.ct, SIMPLIFY = F)

RLEplotsT23scm <- mapply(RLEplot, names= ordnames, logy=normT2scm,samples= map(ys.ct,'samples'), SIMPLIFY = F)

HistsT23scm <- mapply(histT,y=ys.ct,w=map(ruv3T2scm,"W"),topDPC=topDPC,map(ys.ct,'samples'),SIMPLIFY = F)
  
#LFCplotsscm <- mapply(function (x,y) logfcs(DPC=x, pvals=y$pvals,names=c('True','RUVIII')),x=trueLFC, y=HistsT23scm, SIMPLIFY=F)

 
```


### Normalisation {.tabset}


```{r ruv3plots , results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n#### ',celltypes[i], ' {-}\n')))
  
  plots <- list( PCAplotsT23scm[[i]], SilplotsT23scm[[i]], RLEplotsT23scm[[i]])
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=1,ncol=3, common.legend = T, legend="bottom"))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```

### DEA {.tabset}


```{r DE3plots , results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n#### ',celltypes[i], ' {-}\n')))
  
  plots <- list( HistsT23scm[[i]]$hist,HistsT23scm[[i]]$volcano, HistsT23scm[[i]]$volcano_int#,LFCplotsscm[[i]]
                 )
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=2,ncol=2))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```

```{r export ruv3 T4, include=FALSE}

PCAplot(pca= pcas[[3]], y=ys.ct[[3]], avg_sw2 = avgs_sw2[[3]] ,avg_swp = avgs_swp[[3]] ) 
ggsave(paste0(path,'/paper/graphs/studycase/T4PCA3SCm.png'),height = 5, width = 7)

RLEplotsT23scm[[3]]
ggsave(paste0(path,'/paper/graphs/studycase/T4RLE3SCm_scMerge.png'),height = 5, width = 7)

HistsT23scm[[3]]$hist
ggsave(paste0(path,'/paper/graphs/studycase/T4hist3SCm_scMerge.png'),height = 5, width = 7)

```

## Common DEG {.tabset}

<!-- ### Code -->
```{r rank comp}

DEGs <- function(x){
  
  rownames(x$pvals)[x$pvals$adj.P.Val< 0.01]
}
    

DegsUQ <- lapply(Hists,DEGs)
DegsB <- lapply(HistsB ,DEGs)
Degs3 <- lapply(HistsT23,DEGs)
Degs3scm <- lapply(HistsT23scm,DEGs)
ISGsignature <- read.table("/domino/datasets/local/RUV/ISGsignature.txt", quote="\"", comment.char="",header = T)

ISGsignature <- unlist(ISGsignature, use.names = F)
DEGsISG <- lapply(1:8,function(x) ISGsignature)

Degstoplot <- mapply(function(v,w,x,y,s) list(ISG=v,UQ=w,Batch=x,RUVIIIgb=y,
                                                RUVIIIscM=s),v=DEGsISG, w=DegsUQ,x=DegsB,y=Degs3, s=Degs3scm, SIMPLIFY = F)


upsetPlots <- lapply(Degstoplot[1:7], function(x) upset(fromList(x),order.by = "freq") )

degs_methods <- list_transpose(Degstoplot)

ISGdetect <- sapply(degs_methods,function (x) rowSums(sapply(x, function (y) ISGsignature %in% y)))


rownames(ISGdetect)<- ISGsignature
ISGdetect <- as.data.frame(ISGdetect)
ISGdetect$rank <- 1:100
ISGmissRUVIII <- dplyr::select(ISGdetect,-ISG)%>%filter(UQ==0|Batch==0|RUVIIIgb==0|RUVIIIscM==0)


```


```{r upsetplots , results='asis'}

for (i in 1:(length(celltypes)-1)){
  
  cat(knitr::knit_expand(text=paste0('\n### ',celltypes[i], ' {-}\n')))
 
  print(upsetPlots[[i]] )

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```


### Signauture genes not detected

```{r ISGmissRUVIII , results='asis'}

print(knitr::kable(ISGmissRUVIII))

```


```{r t4 upset, include=FALSE}



jpeg(file = paste0(path,'/paper/graphs/studycase/T4upsetSCmRUVIII_scMerge.png'),height = 500, width = 700)
upsetPlots[[3]]
dev.off()



```


## Gene enrichment RUVIII {.tabset}


```{r gsea}

all_gene_sets <- msigdbr(species = "Homo sapiens",category = "C5", subcategory='BP')
genes_padj <- function (pvaltab, q=0.01){
  genes_padj <- rownames(pvaltab[pvaltab$adj.P.Val<0.01,])
  return(genes_padj)
  
}

gseasUQ <- mapply(function(x,y) gsea(genes_padj(x[["pvals"]]),background = row.names(y),geneSets=all_gene_sets,  minSize=5, name=NULL,n=1000),x=Hists, y=ys.ct, SIMPLIFY=F)

gseasBatch <- mapply(function(x,y) gsea(genes_padj(x[["pvals"]]),background = row.names(y),geneSets=all_gene_sets,  minSize=5, name=NULL,n=1000),x=HistsB, y=ys.ct, SIMPLIFY=F)

gseasRUVIIIbg <- mapply(function(x,y) gsea(genes_padj(x[["pvals"]]),background = row.names(y),geneSets=all_gene_sets,  minSize=5, name=NULL,n=1000),x=HistsT23, y=ys.ct, SIMPLIFY=F)


gseasRUVIIIsc <- mapply(function(x,y) gsea(genes_padj(x[["pvals"]]),background = row.names(y),geneSets=all_gene_sets,  minSize=5, name=NULL,n=1000),x=HistsT23scm, y=ys.ct, SIMPLIFY=F)


ISGgseas <-  mapply(function(y) gsea(ISGsignature,background = row.names(y),geneSets=all_gene_sets,  minSize=5, name=NULL,n=1000), y=ys.ct, SIMPLIFY=F)


```

```{r comp}

gseaGenes <- function (gseas){
  gseas[gseas$padj<0.01,]$geneSet
}


estoplot <- mapply(function(v,w,x,y,s) list(ISG=gseaGenes(v),UQ=gseaGenes(w),Batch=gseaGenes(x),RUVIIIgb=gseaGenes(y),
                                                RUVIIIscM=gseaGenes(s)),v=ISGgseas, w=gseasUQ ,x=gseasBatch,y=gseasRUVIIIbg,
                     s=gseasRUVIIIsc, SIMPLIFY = F)

upsetPlotses <- lapply(estoplot[1:7], function(x) upset(fromList(x),order.by = "freq") )

geas_methods <- list_transpose(estoplot)
ISGdetect <- sapply(geas_methods,function (x) rowSums(sapply(x, function (y) gseaGenes(ISGgseas[[1]]) %in% y)))

rownames(ISGdetect)<- gseaGenes(ISGgseas[[1]])
ISGdetect <- as.data.frame(ISGdetect)
ISGdetect$rank <- 1:length(gseaGenes(ISGgseas[[1]]))
ISGmissgsRUVIII <- dplyr::select(ISGdetect,-ISG)%>%filter(UQ==0|Batch==0|RUVIIIgb==0|RUVIIIscM==0)



```

```{r upset gseas, results='asis'}


for (i in 1:(length(celltypes)-1)){

  cat(knitr::knit_expand(text=paste0('\n### ',celltypes[i], ' {-}\n')))
  
  print(upsetPlotses[[i]])
  
  cat(knitr::knit_expand(text=paste0("\n\n")))

}




```


```{r export RUVIII gsea, include=FALSE}


for(i in c(3,4,6,7)){
 jpeg(file = paste0(path,'/paper/graphs/studycase/',celltypes[i],'GSEAupsetSCmRUVIII.png'),height = 500, width = 700)
print(upsetPlotses[[i]])

dev.off() 
}


```


## Different results in RUVIII GSEA {.tabset}

```{r ruvIII diffs , results='asis'}
gseas_methods <- list_transpose(estoplot)
onlygb <- mapply(setdiff,x=gseas_methods$RUVIIIgb,y= gseas_methods$RUVIIIscM)
onlysc <- mapply(setdiff,y=gseas_methods$RUVIIIgb,x= gseas_methods$RUVIIIscM)
diffs <- mapply( function (x,y) data.frame(GOset = c(x,y))%>% mutate(RUVIIIGB= as.numeric(GOset %in% x) , RUVIIIscm= as.numeric(GOset %in% y)), x=onlygb,y=onlysc, SIMPLIFY = FALSE)

lapply(diffs, nrow)

for (i in 1:(length(celltypes)-1)){

  cat(knitr::knit_expand(text=paste0('\n### ',celltypes[i], ' {.tabset}\n')))
  print(knitr::kable(diffs[[i]]))
  cat(knitr::knit_expand(text=paste0("\n\n")))
}

```




## RUVIII PBPS{.tabset}

<!-- ### Code -->

```{r ruv3pbps, include=FALSE}
k=3

psys.ct <- lapply(celltypes,split.ct,PBC=PBPSC)
pslogys.ct <- lapply(psys.ct,function(x)voom(x,plot=F)$E)
psMcts.ct <- lapply (psys.ct, function(x){
  Mct <- replicate.matrix(x$samples$ind_cov)
  rownames(Mct) <- x$samples$ind_cov
  Mct
})

ruv3ps <- mapply(RUVIIIW, Y = lapply(pslogys.ct,t), M=psMcts.ct , ctl =  lapply(psys.ct, function (x) rep(T, nrow(x))), k= as.list(rep(k,length(celltypes))), return.info=as.list(rep(T,length(celltypes))), SIMPLIFY=F)


normpsfull <- lapply(ruv3ps, function (x) t(x$newY))

hvg <- lapply(normpsfull,stvar)
pcas <- mapply(calculatePCA, x= normpsfull, ncomponents=rep(10,length(logys.ct)), subset_row=hvg, SIMPLIFY = F)
PCAplotspsfull <- mapply(PCAplot, pca= pcas, y=psys.ct, SIMPLIFY = F)

orig.s <- lapply(psys.ct, function(x) x$samples$sample_cell[x$samples$pbps==0])
normps <- mapply(function (x,y) t(x$newY[y,]), x=ruv3ps,y=orig.s, SIMPLIFY = F)

hvg <- lapply(normps,stvar)
pcas <- mapply(calculatePCA, x= normps, ncomponents=rep(10,length(logys.ct)), subset_row=hvg, SIMPLIFY = F)

sil2s <- mapply(silhouette,x=pcos,dist=lapply(pcas,function(x) dist(x[,1:2],"euclidean") ), SIMPLIFY = F)
silps <- mapply(silhouette,x=pcos,dist=lapply(pcas,dist, "manhattan") ,SIMPLIFY = F)

avgs_sw2 <- lapply(sil2s, function (x) round(mean(x[,3]),2))
avgs_swp <- lapply(silps, function (x) round(mean(x[,3]),2))

ordnames <- lapply(ys.ct, function (x) arrange(x$samples,Processing_Cohort)$'sample_cell') 

Silplotsps<- lapply(silps,Silplot)

PCAplotsps <- mapply(PCAplot, pca= pcas, y=ys.ct, SIMPLIFY = F)

RLEplotsps <- mapply(RLEplot, names= ordnames, logy=normps,samples= map(ys.ct,'samples'), SIMPLIFY = F)

W <- mapply(function(x,y) x$W[y,], x=ruv3ps,y=orig.s, SIMPLIFY=F)

Histsps <- mapply(histT,y=ys.ct,w=W,topDPC=topDPC,map(ys.ct,'samples'),SIMPLIFY = F)
 
  
```


### Normalisation {.tabset}

```{r ruvpbps plots , results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n#### ',celltypes[i], ' {-}\n')))
  
  plots <- list(PCAplotspsfull[[i]], Silplotsps[[i]], PCAplotsps[[i]], RLEplotsps[[i]])
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=2,ncol=2, common.legend = T, legend="bottom"))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```

### DEA {.tabset}

```{r DEpbpsplots , results='asis'}

for (i in 1:length(celltypes)){
  
  cat(knitr::knit_expand(text=paste0('\n#### ',celltypes[i], ' {-}\n')))
  
  plots <- list( Histsps[[i]]$hist,Histsps[[i]]$volcano,Histsps[[i]]$volcano_int#,LFCplots[[i]]
                 )
 
  print(ggpubr::ggarrange(plotlist=plots,nrow=2,ncol=2))

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```

```{r export pbps T4, include=FALSE}

PCAplot(pca= pcas[[3]], y=ys.ct[[3]], avg_sw2 = avgs_sw2[[3]] ,avg_swp = avgs_swp[[3]] ) 
ggsave(paste0(path,'/paper/graphs/studycase/T4PCApbpsSCm.png'),height = 5, width = 7)

RLEplotsps[[3]]
ggsave(paste0(path,'/paper/graphs/studycase/T4RLEpbpsSCm.png'),height = 5, width = 7)

Histsps[[3]]$hist
ggsave(paste0(path,'/paper/graphs/studycase/T4histpbpsSCm.png'),height = 5, width = 7)


for (i in 1:8){
  Silplotsps[[i]]
ggsave(paste0(path,'/paper/graphs/studycase/',celltypes[i],'silpbpsSCm.png'),height = 5, width = 7)
  
}

```


## Common DEG PBPS{.tabset}

<!-- ### Code -->
```{r rank comp pbps}

DEGs <- function(x){
  
  rownames(x$pvals)[x$pvals$adj.P.Val< 0.01]
}
    

DegsUQ <- lapply(Hists,DEGs)
DegsB <- lapply(HistsB ,DEGs)
DegsPS <- lapply(Histsps,DEGs)


Degstoplot <- mapply(function(v,w,x,
                              z) list(ISG=v,UQ=w,Batch=x, 
                                                RUVIIIpbps=z),v=DEGsISG, w=DegsUQ,x=DegsB,z=DegsPS, SIMPLIFY = F)

upsetPlots <- lapply(Degstoplot[1:7], function(x) upset(fromList(x),order.by = "freq") )

degs_methods <- list_transpose(Degstoplot)

ISGdetect <- sapply(degs_methods,function (x) rowSums(sapply(x, function (y) ISGsignature %in% y)))

rownames(ISGdetect)<- ISGsignature
ISGdetect <- as.data.frame(ISGdetect)
ISGdetect$rank <- 1:100

ISGmissPBPS <- dplyr::select(ISGdetect,-1)%>%filter(UQ==0|Batch==0|RUVIIIpbps==0)

```


```{r upsetplots pbps , results='asis'}

for (i in 1:(length(celltypes)-1)){
  
  cat(knitr::knit_expand(text=paste0('\n### ',celltypes[i], ' {-}\n')))
 
  print(upsetPlots[[i]] )

  cat(knitr::knit_expand(text=paste0("\n\n")))

}
```

### Signauture genes not detected

```{r ISGmissRUVpbps , results='asis'}

print(knitr::kable(ISGmissPBPS))

```



## Gene enrichment {.tabset}


```{r gsea pb}

all_gene_sets <- msigdbr(species = "Homo sapiens",category = "C5", subcategory='BP')
genes_padj <- function (pvaltab, q=0.01){
  genes_padj <- rownames(pvaltab[pvaltab$adj.P.Val<0.01,])
  return(genes_padj)
  
}

gseasUQ <- mapply(function(x,y) gsea(genes_padj(x[["pvals"]]),background = row.names(y),geneSets=all_gene_sets,  minSize=5, name=NULL,n=1000),x=Hists, y=ys.ct, SIMPLIFY=F)

gseasBatch <- mapply(function(x,y) gsea(genes_padj(x[["pvals"]]),background = row.names(y),geneSets=all_gene_sets,  minSize=5, name=NULL,n=1000),x=HistsB, y=ys.ct, SIMPLIFY=F)

gseasRUVIIIpbps <- mapply(function(x,y) gsea(genes_padj(x[["pvals"]]),background = row.names(y),geneSets=all_gene_sets,  minSize=5, name=NULL,n=1000),x=Histsps, y=ys.ct, SIMPLIFY=F)

ISGgseas <-  mapply(function(y) gsea(ISGsignature,background = row.names(y),geneSets=all_gene_sets,  minSize=5, name=NULL,n=1000), y=ys.ct, SIMPLIFY=F)


```

```{r comp pb}

gseaGenes <- function (gseas){
  gseas[gseas$padj<0.1,]$geneSet
}


estoplot <- mapply(function(v,w,x,y) list(ISG=gseaGenes(v),UQ=gseaGenes(w),Batch=gseaGenes(x),RUVIIIpbps=gseaGenes(y)),v=ISGgseas, w=gseasUQ ,x=gseasBatch,y=gseasRUVIIIpbps, SIMPLIFY = F)

upsetPlotses <- lapply(estoplot[1:7], function(x) upset(fromList(x),order.by = "freq") )

gseas_methods <- list_transpose(estoplot)
ISGdetect <- sapply(gseas_methods,function (x) rowSums(sapply(x, function (y) gseaGenes(ISGgseas[[1]]) %in% y)))

rownames(ISGdetect)<- gseaGenes(ISGgseas[[1]])
ISGdetect <- as.data.frame(ISGdetect)
ISGdetect$rank <- 1:length(gseaGenes(ISGgseas[[1]]))
ISGmiss <- dplyr::select(ISGdetect,-ISG)%>%filter(UQ==0|Batch==0|RUVIIIpbps==0)


UQmiss1 <- sapply(gseas_methods,function(w) mapply(function(x,y) x[!x %in%  y],x=gseas_methods[['UQ']], y =w ),simplify = F)

gseasAllUQ <- unique(unlist(gseas_methods[['UQ']]))

gseasUQdetect <- sapply(gseas_methods,function (x) rowSums(sapply(x, function (y) gseasAllUQ  %in% y)))

rownames(gseasUQdetect)<-gseasAllUQ 
gseasUQdetect <- as.data.frame(gseasUQdetect)
gseasUQmiss <- dplyr::select(gseasUQdetect,-ISG)%>%filter(Batch==0,RUVIIIpbps==0)

UQexport <- all_gene_sets[all_gene_sets$gs_name%in% rownames(gseasUQmiss),'gs_exact_source']%>%unique()

#write.table(UQexport,'UQonlyGSEApbpsCS.txt',quote=F, row.names = F) 

```

```{r upset gseas pb, results='asis'}


for (i in 1:(length(celltypes)-1)){

  cat(knitr::knit_expand(text=paste0('\n### ',celltypes[i], ' {-}\n')))
  
  print(upsetPlotses[[i]])
  


}

```

```{r pbps gsea upset, include=FALSE}


for(i in c(3,4,6,7)){
 jpeg(file = paste0(path,'/paper/graphs/studycase/',celltypes[i],'GSEAupsetSCmPBPS.png'),height = 500, width = 700)
print(upsetPlotses[[i]])

dev.off() 
}




```


## Different results in GSEA {.tabset}

```{r ruvpbps diffs , results='asis'}
gseas_methods <- list_transpose(estoplot)
onlyBatch <- mapply(setdiff,x=gseas_methods$Batch,y= gseas_methods$RUVIIIpbps)
onlypbps <- mapply(setdiff,y=gseas_methods$Batch,x= gseas_methods$RUVIIIpbps)

diffs <- mapply( function (x,y) data.frame(GOset = c(x,y))%>% mutate(Batch= as.numeric(GOset %in% x) , PBPS= as.numeric(GOset %in% y)), x=onlyBatch,y=onlypbps, SIMPLIFY = FALSE)


for (i in 1:(length(celltypes)-1)){

  cat(knitr::knit_expand(text=paste0('\n### ',celltypes[i], ' {.tabset}\n')))
  print(knitr::kable(diffs[[i]]))
  cat(knitr::knit_expand(text=paste0("\n\n")))
  
}

```

